Measurement and Probability in Fuzzy Quantum Theories

### 2016-12-10 prop.tex
prop.tex \section{Introduction} Add citations...

This is a \emph{theoretical} investigation of \emph{experimental}
physics using \emph{computational} methods. All experiments and computations
are processes bounded in space, time, energy, and other resources~\cite{Jaeger2007,Li2009}.
Yet, for centuries, the mathematical formalization of such processes
has been founded on the infinitely precise real or complex numbers.
Our purpose here is exploit the consequences of replacing infinitely-specified
quantum probabilities by finite number of intervals used in interval-valued
probability measures; in particular, we show that a quantum interval-valued
probability measure may not be induced by a infinitely-precise state,
but it might be more and more likely to be identified to a particular
state when the measurement resources increase.

Indeed, almost every description of quantum mechanics, quantum computation,
or quantum experiments refers to entities such as $\rme$, $\pi$,
$\sqrt{2}$, etc. From a computational perspective, such numbers do
not exist in their entirety ``for free~\cite{Kent1999,CliftonKent2000}.''
For example, the state of the art algorithms for computing the $n$th
binary digit of~$\pi$ require on the order of $O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, simply
referring to the $n$th digit of $\pi$ requires more and more resources
as $n$ gets larger. Taking such resource bounds into consideration
is what founded computer science as a discipline and is crucial for
understanding the very nature of computation and, following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, for understanding the
very nature of physical processes.

We propose to revisit quantum mechanics, quantum information, and
quantum computation from this resource-aware perspective. Our initial
results in that domain showed how subtle the issues can
be~\cite{usat,geometry2013,DQT2014}: a straightforward replacement of
the complex numbers by a finite field yields a variant of quantum
mechanics in which computationally hard problems like UNIQUE-SAT
(which decides whether a given Boolean formula is unsatisfiable or has
exactly one satisfying assignment) can be deterministically solved in
constant time. To eliminate such unrealistic theories requires
delicate analysis of the structure of the Hilbert space, the process
of observation, and the notion of probability teasing apart their
reliance on the infinitely precise real
numbers~\cite{geometry2013,DQT2014}.

In this proposal, our aim is to
shift focus from the infinitely-specified but not directly observable
quantum states, to observable measurable properties of quantum systems
and their probabilities. Furthermore, we insist that our theories of
measurement and probability only refer to finitely communicable
evidence within feasible computational bounds. It follows that states,
observations, and probabilities all become ``fuzzy'', i.e., specified
by intervals of confidence that can only increase in precision if the
available resources increase proportionally. Our notion of ``fuzzy
quantum mechanics'' is related to existing
work~\cite{GranikCaulfield1996,Pykacz2013,SNL2009,Gudder2005,aerts1993physical}
but, as will be explained in more detail, is distinguished by its
unique computational character.

We will begin by reviewing existing work that recasts classical
probability spaces in a resource-aware setting and move to our
proposal which, briefly speaking, aims at recasting quantum
probability and hence quantum measurement to a corresponding
resource-aware setting.

In particular, the objectives to be achieved include:
\begin{itemize}
\item Develop a measurement framework based on quantum interval-valued
  probabilities measures.
\item Assess the validity of fundamental theorems of quantum
  mechanics, such as Gleason, Bell, and Kochen-Specker, in a quantum
  formulation stripped away of infinite resources. We will aim to
  provide a definite answer to the Meyer-Mermin
  debate~\cite{PhysRevLett.83.3751, Mermin1999, BarrettKent2004} on
  the impact of finite precision measurements on the relevance of the
  Kochen-Specker
  theorem~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}.
\end{itemize}

@@

If we insist that probabilities cannot be computed to infinite precision
and are bound to be approximations represented by intervals of confidence,
then quantum states themselves can only be discussed within intervals
of confidence. Despite the fact that classically a interval-valued
probability measure could always correspond to a real-valued probability
measure, we found that there may be no ``real'' infinitely-precise
quantum state approximated by a quantum interval-valued probability
measure as a whole. However, if a quantum interval-valued probability
measure is decomposed into pieces, each piece might still be induced
by a quantum state. Moreover, our examples and Gleason's theorem summarized
in table~\ref{table_long} suggests that as the measurement resources
increase, an entire quantum interval-valued probability measure could
more and more likely be identified to a particular state. 
\begin{table}[ht]
\center{%
\begin{tabular}{c|cccc}
Measurement resources & Lowest & \multicolumn{2}{c}{$\longleftrightarrow$} & Highest\tabularnewline
\hline 
Count of $\mathscr{I}$ & $3$ & $4$ & \ldots{} & $\infty$\tabularnewline
$\sup_{\left[l,r\right]\in\mathscr{I}}\left|r-l\right|$ & $1$ & $\frac{1}{2}$ & \ldots{} & $0$\tabularnewline
Count of $\mathrm{core}\left(\bar{\mu}\right)$ & 0 & 0 &  & 1\tabularnewline
Count of states might correspond to $\bar{\mu}$ & $3$ & $2$ & \ldots{} & $1$\tabularnewline
\hline 
How precise we could identify a state? & Coarse & \multicolumn{2}{c}{$\longleftrightarrow$} & Precise\tabularnewline
\end{tabular}}\caption{Relation between measurement resources and interval-valued probability
measures}
\label{table_long}
\end{table}
Additional work is in progress on the following research questions:
\begin{description}
\item [{Research Question.}] Confirm that there is always a quantum interval-valued
probability measure with an empty-core if the length of the intervals
is bigger than zero. 
\item [{Research Question.}] Find a general method to determine how a
quantum interval-valued probability measure is close to the Born rule. 
\item [{Research Question.}] Confirm that as the number and precision
of the intervals increase the quantum interval-valued probability
measure converges to the measure induced by the Born rule. 
\item [{Research Question.}] Investigate the status of the theorems of
Gleason, Bell~\cite{BellBook1987,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and Kochen-Specker for quantum interval-valued probability measures.
\end{description}
Although we discussed how a quantum interval-valued probability measure
could be induced from a real quantum state, our investigation leaves
an open question of whether there exists a ``real'' entity that
exists independently of measurements and probabilities. The possibility
of no ``real'' underlying state is consistent with the elegantly
recent work on Quantum Bayesianism or QBism~\cite{Fuchs2010,VonBaeyer2016,Fuchs2012},
which suggests that the quantum state is more like an interactive
system in computer science parlance. In another word, the quantum
state is subjective: each observer has a different view of the quantum
system that is consistent with their previous observations and that
allows that observer, independently of other observers, to assign
beliefs (i.e., probabilities) to possible future interactions.

>>>>

This is a \emph{theoretical} investigation of \emph{experimental}
physics using \emph{computational} methods. All experiments and computations
are processes bounded in space, time, energy, and other resources~\cite{Jaeger2007}.
Yet, for centuries, the mathematical formalization of such processes
has been founded on the infinitely precise real or complex numbers.
Our purpose here is to exploit the consequences of replacing infinitely-specified
quantum probabilities by finite number of intervals used in interval-valued
probability measures; in particular, we show that a quantum interval-valued
probability measure may not be induced by a infinitely-precise state,
but it might be more and more likely to be identified to a particular
state when the measurement resources increase.

Indeed, almost every description of quantum mechanics, quantum computation,
or quantum experiments refers to entities such as $\rme$, $\pi$,
$\sqrt{2}$, etc (see, e.g., \cite{Redhead1987-REDINA,544199,Mermin2007}).
From a computational perspective, such numbers do not exist in their
entirety ``for free~\cite{Kent1999,CliftonKent2000}.'' For example,
the state of the art algorithms for computing the $n$th binary digit
of~$\pi$ require on the order of $O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, simply
referring to the $n$th digit of $\pi$ requires more and more resources
as $n$ gets larger. Taking such resource bounds into consideration
is what founded computer science as a discipline and is crucial for
understanding the very nature of computation and, following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, for understanding the
very nature of physical processes.

We have been revisiting quantum mechanics, quantum information, and
quantum computation from this resource-aware perspective. Our initial
results in that domain showed how subtle the issues can be~\cite{usat,geometry2013,DQT2014}:
a straightforward replacement of the complex numbers by a finite field
yields a variant of quantum mechanics in which computationally hard
problems like UNIQUE-SAT (which decides whether a given Boolean formula
is unsatisfiable or has exactly one satisfying assignment~\cite{Valiant198685,Papadimitriou1993,AroraBarak2009})
can be deterministically solved in constant time. To eliminate such
unrealistic theories requires delicate analysis of the structure of
the Hilbert space, the process of observation, and the notion of probability
teasing apart their reliance on the infinitely precise real numbers~\cite{geometry2013,DQT2014}.

In this paper, we shift focus from the infinitely-specified but not
directly observable quantum states, to observable measurable properties
of quantum systems and their probabilities. Furthermore, we insist
that our theories of measurement and probability only refer to finitely
communicable evidence within feasible computational bounds. It follows
that states, observations, and probabilities all become ``fuzzy'',
i.e., specified by intervals of confidence that can only increase
in precision if the available resources increase proportionally. Our
notion of ``fuzzy quantum mechanics'' is related to existing work~\cite{GranikCaulfield1996,Pykacz2013,SNL2009,Gudder2005,aerts1993physical}
but, as will be explained in more detail, is distinguished by its
unique computational character.

We will begin by reviewing existing work that recasts classical probability
spaces in a resource-aware setting and move to aim at recasting quantum
probability and hence quantum measurement to a corresponding resource-aware
setting. In particular, we develop a measurement framework based on
quantum interval-valued probabilities measures. 
\begin{itemize}
\item Assess the validity of fundamental theorems of quantum mechanics,
such as Gleason~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum},
Bell~\cite{BellBook1987,Redhead1987-REDINA,peres1995quantum,Jaeger2007},
and Kochen-Specker~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007},
in a quantum formulation stripped away of infinite resources. We will
aim to provide a definite answer to the Meyer-Mermin debate~\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004}
on the impact of finite precision measurements on the relevance of
the Kochen-Specker theorem. 
\end{itemize}

@@

If we insist that probabilities cannot be computed to infinite precision
and are bound to be approximations represented by intervals of confidence,
then quantum states themselves can only be discussed within intervals
of confidence. Despite the fact that classically a interval-valued
probability measure could always correspond to a real-valued probability
measure, we found that there may be no ``real'' infinitely-precise
quantum state approximated by a quantum interval-valued probability
measure as a whole. However, if a quantum interval-valued probability
measure is decomposed into pieces, each piece might still be induced
by a quantum state. Moreover, our examples and Gleason's theorem summarized
in table~\ref{table_long} suggests that as the measurement resources
increase, an entire quantum interval-valued probability measure could
more and more likely be identified to a particular state. 
\begin{table}[ht]
\center{%
\begin{tabular}{c|cccc}
Measurement resources & Lowest & \multicolumn{2}{c}{$\longleftrightarrow$} & Highest\tabularnewline
\hline 
Count of $\mathscr{I}$ & $3$ & $4$ & \ldots{} & $\infty$\tabularnewline
$\sup_{\left[l,r\right]\in\mathscr{I}}\left|r-l\right|$ & $1$ & $\frac{1}{2}$ & \ldots{} & $0$\tabularnewline
Count of $\mathrm{core}\left(\bar{\mu}\right)$ & 0 & 0 &  & 1\tabularnewline
Count of states might correspond to $\bar{\mu}$ & $3$ & $2$ & \ldots{} & $1$\tabularnewline
\hline 
How precise we could identify a state? & Coarse & \multicolumn{2}{c}{$\longleftrightarrow$} & Precise\tabularnewline
\end{tabular}}\caption{Relation between measurement resources and interval-valued probability
measures}
\label{table_long}
\end{table}
Additional work is in progress on the following research questions:
\begin{description}
\item [{Research Question.}] Confirm that there is always a quantum interval-valued
probability measure with an empty-core if the length of the intervals
is bigger than zero. 
\item [{Research Question.}] Find a general method to determine how a
quantum interval-valued probability measure is close to the Born rule. 
\item [{Research Question.}] Confirm that as the number and precision
of the intervals increase the quantum interval-valued probability
measure converges to the measure induced by the Born rule. 
\item [{Research Question.}] Investigate the status of the theorems of
Gleason, Bell, and Kochen-Specker for quantum interval-valued probability
measures.
\end{description}
Although we discussed how a quantum interval-valued probability measure
could be induced from a real quantum state, our investigation leaves
an open question of whether there exists a ``real'' entity that
exists independently of measurements and probabilities. The possibility
of no ``real'' underlying state is consistent with the elegantly
recent work on Quantum Bayesianism or QBism~\cite{Fuchs2010,VonBaeyer2016,Fuchs2012},
which suggests that the quantum state is more like an interactive
system in computer science parlance. In another word, the quantum
state is subjective: each observer has a different view of the quantum
system that is consistent with their previous observations and that
allows that observer, independently of other observers, to assign
beliefs (i.e., probabilities) to possible future interactions.
